Project Description -- Erosion, Transport, Anomalous Diffusion, and Collapse in Granular and Porous Media

### 2019-11-13 docs/proposals/NSF/ProjDesc/ProjDesc.tex
NM

An understanding of these filtration and contamination processes relies on the complex phenomenon of transport in porous media. Further, the action of continually flowing groundwater can alter medium properties through erosive and transportive processes. Thus, it becomes necessary to understand transport through complex media whose properties change {\em dynamically} in response to the intervening fluid flow. Such effects become most prominent during rapid events, like the gravitational collapse of a sinkhole~\cite{sandhu2018fate}, during which medium properties can change suddenly and dramatically. Though less obvious, the accumulation of slower process, such as mechanical or chemical erosion, can also substantially alter medium properties.

This proposal aims to develop a suite of computational tools geared towards developing a deeper understanding of the physical processes by which groundwater flow can alter porous-media properties and to characterize the associated changes in transport within the evolving medium, for example the anomalous dispersion of tracers. A recognized complexity of naturally occurring media is the heterogeneity and anisotropy of macroscopic properties such as permeability and mechanical strength~\cite{lin2018randomization}. These properties can change over short timescales, for example sinkhole collapse or hydraulic fracturing, or over long timescales, for example the slow erosion or dissolution of solid material. Potential exists for feedback between these timescales: for example the slow wearing process reaching a critical threshold beyond which catastrophic events become more likely.  Physical mechanisms that can dynamically alter porous-medium properties include: 
\begin{itemize}[noitemsep]
\item Erosion of solid material due to fluid-mechanical stresses or chemical dissolution.
\item Transport of grains due to intervening fluid flow and due to sedimentation.
\item Compaction of the medium due to overlying loads.
\item Gravitational collapse of a medium due to loads exceeding a threshold.
\end{itemize}
The above are roughly ordered according to timescale, from the slowest acting to the fastest. While many of these mechanisms have been studied individually in various contexts and to varying levels of rigor, very little is known about their combined effect on the dynamic evolution of a porous medium. Since we aim to build a quantitative model from the ground up, we will first focus on the slowest acting mechanisms, as these ultimately set the stage for the more rapid events like sinkhole collapse.

@@

\item Additionally effects, such as sedimentation, transport, and compaction constantly work to reconfigure the medium as its constituents and boundaries degrade.
\end{itemize}
These challenges require innovative computational techniques, some of which have been developed recently~\cite{Quaife2018} and others that await resolution. An ultimate goal of this proposal is to achieve high-fidelity simulations of $\mathcal{O}(100-1000)$ mobile bodies undergoing simultaneous erosion, transport, sedimentation, and compaction, with the computational efficiency required to run large ensembles for statistical analysis. This is an highly active area of research, as closely related laboratory experiments on erosion in colloidal systems are already being explored in the Datta Lab at Princeton~\cite{bizmark2019multiscale}. These experiments a great opportunity for comparison against our numerical simulations. Furthermore, we propose performing a set of experiments conducted at the nearby the Geophysical Fluid Dynamics Institute, which would help to guide, calibrate, and validate the computational methods.

@@

While erosion and dissolution lead to gradual changes in a medium, other events like sinkhole collapse can occur suddenly and catastrophically.  Most likely, all of the mechanisms listed above---erosion, dissolution, transport, compaction, and gravitational collapse---contribute to sinkhole formation, with the slowest acting mechanisms creating conditions favorable for the final act of gravity. Sinkhole formation has been explored with preliminary experiments run at the Geophysical Fluid Dynamics Institute (GFDI) at Florida State University~\cite{tao2014experimental}. These scaled-down experiments offer the ability to probe various mechanisms behind sinkhole formation, for example a difference in hydraulic head between the confined and unconfined aquifers. An ultimate goal of this project is to merge these two lines of inquiry to gain a deeper understanding the influence of mechanically or chemically induced erosion on gravitational collapse.  Ultimately, we envision the high-fidelity erosion simulations to inform coarse-grained models for: (i) transport and dispersion; and (ii) dynamic evolution of macroscopic medium properties. These parameterized models could then be combined with well-developed multi-phase models~\cite{Imma2019, eastham2019multiphase} to be used for the quantitative description and prediction of sinkhole collapse.

In the following, we first describe the underlying governing equations, then detail the computational challenges and proposed work on simulating fluid-mechanical erosion, characterizing transport and dispersion, and modeling gravitational interactions.

%---------------------------------------------------------------------------------------------------%
\subsection{Governing equations}
At the finest scale, the fluid flow intervening between individual grains of a porous medium is governed by the incompressible Stokes equations
\begin{equation}
\label{StokesEq}
\begin{split}
\mu \Delta \uu = \grad p,	&\hspace{20pt} \xx \in \Omega, \\
\grad \cdot \uu = 0, 		&\hspace{20pt} \xx \in \Omega,
\end{split}
\end{equation}
where $\uu$ is the velocity field, $p$ the pressure field, and $\mu$ the fluid viscosity. Linear elliptic PDEs, such as the Stokes equations, can be reformulated as a boundary integral equation (BIE).  BIEs have several numerical advantages including a dimension reduction, automatic mass conservation, and high-order accuracy. An equivalent BIE formulation of the Stokes equations in Eq.~\eqref{StokesEq} is
\begin{align}
  \label{eqn:BIE}
  \ff(\xx) = -\frac{1}{2}\eeta(\xx) + \frac{1}{\pi} \int_{\bd\Omega}
    \frac{\rr \cdot \nn}{\rho^2} \frac{\rr \otimes \rr}{\rho^2}
    \eeta(\yy) ds_\yy, \quad \xx \in \bd\Omega
\end{align}
where $\ff$ is a Dirichlet boundary condition, $\rr = \xx - \yy$, $\rho = \|\rr\|$, and $\nn$ is the outward unit normal of $\bd\Omega$.  Once~\eqref{eqn:BIE} is solved for the density function $\eeta$, important quantities such as the velocity, pressure, vorticity, and deformation tensor can be computed in the fluid bulk.  For example, the velocity at $\xx \in \Omega$ is 
\begin{align*}
  \uu(\xx) = \frac{1}{\pi} \int_{\bd\Omega}
    \frac{\rr \cdot \nn}{\rho^2} \frac{\rr \otimes \rr}{\rho^2}
    \eeta(\yy) ds_\yy.
\end{align*}

@@

While the most immediate proposed work on computational methods to simulate erosion will utilize~\eqref{StokesEq}, the later stages of this project will also incorporate~\eqref{DarcyEq}, or closely related variants known as multiphase models~\cite{Imma2019, eastham2019multiphase}, in order to develop coarse-grained descriptions for porous medium evolution and the associated dispersion within. These models will rely on statistical analysis of the data produced by the high-fidelity erosion simulations in order to parameterize the effects of erosion, leading to deterministic or stochastic reduced-order models for porous-medium evolution and erosion, combined with other external effects such as gravitationally-induced dynamics.

@@

The creation and reinforcement of channels significantly impacts the macroscopic properties of the porous medium. Fig.~\ref{fig3} shows new, preliminary results on how media properties change in time. First, the left-most plot shows the rate at which the area fraction decreases as the bodies erode. Time is normalized by the value $t_f$ at which all bodies vanish and hence the area fraction is zero. Notice that the rate of area reduction does not change significantly during the simulation. The middle plot shows the resistivity, or inverse permeability of the medium. In the simulations, resistivity is calculated by measuring the total flux and the pressure upstream and downstream, then using Darcy's relation \eqref{DarcyEq} to infer $k$. Naturally, as bodies erode, the resistivity they provide decreases in time. Note, however, that the vertical axis is logarithmic. Hence, the initial linear decrease corresponds to an exponential rate of resistivity reduction. The much more mild decrease in area fraction is not sufficient to account for this exponential rate, and hence the reshaping process, in particular the formation of channels, plays a pivotal role.

@@

An immediate goal of the proposed work is extend the range of physical effects captured in these simulations. A particular challenge of a BIE formulation is developing quadrature for nearly-singular integrands that arise when bodies are close. To address this problem, we have very recently extended a Barycentric quadrature~\cite{abr2014, bar-wu-vee2015}---we have successfully simulated erosion with grains that are initially separated by 1\% of an arclength spacing. This quadrature approach has the advantage that it is a non-intrusive modification of the trapezoid rule, allowing us to achieve an error bound that independent of the grain configuration. Near-term plans are to include flow transport and gravitational sedimentation of the eroding grains. For this extension, contact forces between bodies must be included to eliminate unphysical contact that is inevitable due to discretization errors. Though conceptually straightforward, implementing contact forces in multibody fluid-structure simulations is far from simple in practice, with several competing possible strategies that are a topic of current research.  One approach is to use introduce a potential with short-range repulsion~\cite{}, but such methods often introduce stiffness. Alternatively, a space-time interference volume can e used to guarantee a minimum separation between grains.  In a fluid context, this method was first introduced by Lu et al.~\cite{lu-rah-zor2017}, and extended to allow for much smaller separation distances of rigid bodies by the PI's previous graduate student~\cite{bys-sha-qua2019}. One Ph.D.~student, who will be jointly supervised by both PIs, will simulate sedimentation of eroding bodies by developing algorithms that resolve quadrature error and contact.

@@

@@

Transport through eroded media is a direct consequence of the pore
structure. Qualitative properties of an eroded media, such as channels
of high porosity can be modest in volume fraction, but still be
responsible for a large portion of the flux~\cite{Quaife2018}. To
increase understanding of flow through eroded media, we will simulate
and statistically analyze transport through eroded geometries.

In an eroded geometry $\Omega$, a concentration non-dimensionalized
concentration $c$ is governed by
\begin{align}
  \pd{c}{t} + \uu \cdot \nabla c = \frac{1}{\Pe} \Delta c, 
  \label{eqn:advectionDiffusion}
\end{align}
where $\Pe$ is the dimensionless Peclet number. Here we assume that the
concentration $c$ is affected by the eroded geometry, but the geometry
is not altered further by the concentration. In the infinite Peclet
number limit, the concentration is governed by pure advection and the
characteristics of the concentration field are identical to the
streamlines (Figure~\ref{fig:100tracers}).
\begin{wrapfigure}[]{l}{0.4\textwidth}
  \includegraphics[width=0.4\textwidth]{figs/tort_eulerian100}
  \caption{\label{fig:100tortuosity} The tortuosity of an eroding
  geometry. The tortuosity can be computed using two different formulas
  (red vs.~blue marks). The red square is the tortuosity in
  Figure~\ref{fig:100tracers}. Their is qualitative agreement between
  our numerical simulations and a power-law model (black dashed
  line)~\cite{mat-kha-koz2008}.}
\end{wrapfigure}

The $j^{th}$ trajectories, or characteristics, is governed by
$\dot{\ss}_j(t) = \uu(\ss_j(t))$. Once these trajectories are computed,
using, for example, a high-order Runge-Kutta method, a statistical
analysis be used to describe the transport. Two of the most important
measurements are the tortuosity and the anomalous dispersion, both which
are defined in terms of the length of the trajectories
\begin{align}
  \lambda_j(t) = \int_{0}^{t} \|\ss'(t')\| dt'.
\end{align}
Once $N_p$ trajectories are computed, the tortuosity is the
dimensionless number
\begin{align}
  T = \frac{1}{d} \left(\int_{S} u_1(x_0,y_0)\lambda(y_0) dy_0\, \right)
  \Bigg/ \left(\int_S u_1(x_0,y_0)\, dy_0 \right),
\end{align}
where $S$ is the vertical cross-section at $x=x_0$, $d$ is its length,
$u_1$ is the horizontal velocity, and $\lambda(y_0)$ is the length of
the trajectory starting at $(x_0,y_0)$ and ending at a vertical
cross-section at the other end of the channel. Relationships between the
tortuosity and porosity of a geometry have been developed by other
groups~\cite{kop-kat-tim1996, dud-koz-mat2011, mat-kha-koz2008}, and
preliminary results indicate that such models might be appropriate for
geometries whose porosity is increasing under the process of erosion
(Figure~\ref{fig:100tortuosity}).

@@

As mentioned above, researchers at FSU's Geophysical Fluid Dynamics Institute (GFDI) have investigated the mechanisms behind sinkhole formation with a series of scaled-down laboratory experiments~\cite{tao2014experimental}. The new simulations will compare against this experimental data and guide future experiments.

MORE

\np
%---------------------------------------------------------------------------------------------------%
%% BROADER IMPACTS
%\section{Broader Impacts of the Proposed Work}
% Xiaoming
%\subsection{Broader scientific impact of the project} 
%\subsection{Student training and collaborative work}

>>>>

An understanding of these filtration and contamination processes relies on the complex phenomenon of transport in porous media. Further, the action of continually flowing groundwater can alter medium properties through erosive and transportive processes. Thus, it becomes necessary to understand transport through complex media whose properties change {\em dynamically} in response to the intervening fluid flow. These effects are most noticeable during rapid events, like the gravitational collapse of a sinkhole~\cite{sandhu2018fate}, during which medium properties can change dramatically over short timescales. Though less obvious, the accumulation of slower process, such as mechanical or chemical erosion, can also substantially alter medium properties.

This proposal aims to develop a suite of computational tools geared towards developing a deeper understanding of the physical processes by which groundwater flow can alter porous-media properties and to characterize the associated changes in transport within the evolving medium, for example the anomalous dispersion of tracers. A recognized complexity of naturally occurring media is the heterogeneity and anisotropy of macroscopic properties such as permeability and mechanical strength~\cite{lin2018randomization}. These properties can change over short timescales, for example sinkhole collapse or hydraulic fracturing, or over long timescales, for example the slow erosion or dissolution of solid material. Potential exists for feedback between these timescales: for example the slow wearing process reaching a critical threshold beyond which catastrophic events become more likely.  Physical mechanisms that can alter porous-medium properties include: 
\begin{itemize}[noitemsep]
\item Erosion of solid material due to fluid-mechanical stresses or chemical dissolution.
\item Transport of grains due to intervening fluid flow and due to sedimentation.
\item Compaction of the medium due to overlying loads.
\item Gravitational collapse of a medium due to loads exceeding a threshold.
\end{itemize}
The above are roughly ordered according to timescale, from the slowest acting to the fastest. While many of these mechanisms have been studied individually in various contexts and to varying levels of rigor, very little is known about their combined effect on the dynamic evolution of a porous medium. Since we aim to build a quantitative model from the ground up, we will first focus on the slowest acting mechanisms, as these ultimately set the stage for the more rapid events like sinkhole collapse.

@@

\item Additionally effects, such as sedimentation, transport, and compaction constantly work to reconfigure the medium as its constituents and boundaries degrade.
\end{itemize}
These challenges require innovative computational techniques, some of which have been developed recently~\cite{Quaife2018} and others that await resolution. An ultimate goal of this proposal is to achieve high-fidelity simulations of $\mathcal{O}(100-1000)$ mobile bodies undergoing simultaneous erosion, transport, sedimentation, and compaction, with the computational efficiency required to run large ensembles for statistical analysis. Dynamic erosion of porous media is a highly active area of research, as closely related laboratory experiments on erosion in colloidal systems are already being explored in the Datta Lab at Princeton~\cite{bizmark2019multiscale}. These experiments a great opportunity for comparison against our numerical simulations. Furthermore, we propose performing a set of experiments conducted at the nearby the Geophysical Fluid Dynamics Institute, which would help to guide, calibrate, and validate the computations.

@@

While erosion and dissolution lead to gradual changes in a medium, other events like sinkhole collapse or hydraulic fracturing occur over much shorter timescales.  Most likely, all of the mechanisms listed above---erosion, dissolution, transport, compaction, and gravitational collapse---contribute to sinkhole formation, with the slowest acting mechanisms creating conditions favorable for the final act of gravity. Sinkhole formation has been explored in preliminary experiments conducted at the Geophysical Fluid Dynamics Institute (GFDI) at Florida State University~\cite{tao2014experimental}. These scaled-down experiments offer the ability to probe various mechanisms behind sinkhole formation, for example a difference in hydraulic head between the confined and unconfined aquifers. An ultimate goal of the project proposed here is to merge these two lines of inquiry to gain a deeper understanding of the influence of mechanically or chemically induced erosion on gravitational collapse.  Ultimately, we envision the high-fidelity erosion simulations to inform coarse-grained models for: (i) transport and dispersion; and (ii) dynamic evolution of macroscopic medium properties. These parameterized models could then be combined with well-developed multi-phase models~\cite{Imma2019, eastham2019multiphase} to quantitatively describe and predict sinkhole collapse.

%---------------------------------------------------------------------------------------------------%
\subsection{Governing equations}

We first summarize governing equations underlying the proposed work.
At the finest scale, the fluid flow intervening between individual grains of a porous medium is governed by the incompressible Stokes equations
\begin{equation}
\label{StokesEq}
\begin{split}
\mu \Delta \uu = \grad p,	&\hspace{20pt} \xx \in \Omega, \\
\grad \cdot \uu = 0, 		&\hspace{20pt} \xx \in \Omega,
\end{split}
\end{equation}
where $\uu$ is the velocity field, $p$ the pressure field, and $\mu$ the fluid viscosity. Linear elliptic PDEs, such as the Stokes equations, can be reformulated as a boundary integral equation (BIE).  BIEs have several numerical advantages including a dimension reduction, automatic mass conservation, and high-order accuracy. An equivalent BIE formulation of the Stokes equations in Eq.~\eqref{StokesEq} is
\begin{align}
  \label{eqn:BIE}
  \ff(\xx) = -\frac{1}{2}\eeta(\xx) + \frac{1}{\pi} \int_{\bd\Omega}
    \frac{\rr \cdot \nn}{\rho^2} \frac{\rr \otimes \rr}{\rho^2}
    \eeta(\yy) ds_\yy, \quad \xx \in \bd\Omega
\end{align}
where $\ff$ is a Dirichlet boundary condition, $\rr = \xx - \yy$, $\rho = \|\rr\|$, and $\nn$ is the outward unit normal of $\bd\Omega$.  Once~\eqref{eqn:BIE} is solved for the density function $\eeta$, important quantities such as the velocity, pressure, vorticity, and deformation tensor can be computed in the fluid bulk.  For example, the velocity at $\xx \in \Omega$ is 
\begin{align*}
  \uu(\xx) = \frac{1}{\pi} \int_{\bd\Omega}
    \frac{\rr \cdot \nn}{\rho^2} \frac{\rr \otimes \rr}{\rho^2}
    \eeta(\yy) ds_\yy.
\end{align*}

@@

While the most immediate proposed work on computational methods to simulate erosion will utilize~\eqref{StokesEq}, the later stages of this project will also incorporate~\eqref{DarcyEq}, or closely related variants known as multiphase models~\cite{Imma2019, eastham2019multiphase}, in order to develop coarse-grained descriptions for porous medium evolution and the associated dispersion within. These models will rely on statistical analysis of the data produced by the high-fidelity erosion simulations in order to parameterize the effects of erosion, leading to deterministic or stochastic reduced-order models for porous-medium evolution. 
%and erosion, combined with other external effects such as gravitationally-induced dynamics.

@@

The creation and reinforcement of channels significantly impacts the macroscopic properties of the porous medium. Fig.~\ref{fig3} shows new, preliminary results on how media properties change in time. First, the left-most plot shows the rate at which the area fraction decreases as the bodies erode. 
%Time is normalized by the value $t_f$ at which all bodies vanish and hence the area fraction is zero. 
Notice that the rate of area reduction does not change significantly during the simulation. The middle plot shows the resistivity, or inverse permeability of the medium. In the simulations, resistivity is calculated by measuring the total flux and the pressure upstream and downstream, then using Darcy's relation \eqref{DarcyEq} to infer the horizontal permeability $k_x$. Naturally, as bodies erode, the resistivity they provide decreases in time. Note that, unlike the area plot, the vertical axis is logarithmic and so the straight line observed at early times indicates that the resistivity decreases exponentially. The much more mild decrease in area fraction is not sufficient to account for this exponential rate, and hence the reshaping process, in particular the formation of channels, plays a pivotal role.

@@

An immediate goal of the proposed work is extend the range of physical effects captured in these simulations. A particular challenge of a BIE formulation is developing quadrature for nearly-singular integrands that arise when bodies are close. To address this problem, we have very recently extended a Barycentric quadrature~\cite{abr2014, bar-wu-vee2015} and have successfully simulated erosion with grains that are initially separated by 1\% of an arclength spacing. This quadrature approach has the advantage that it is a non-intrusive modification of the trapezoid rule, allowing us to achieve an error bound that independent of the grain configuration. Near-term plans are to include flow transport and gravitational sedimentation of the eroding grains. For this extension, contact forces between bodies must be included to eliminate unphysical contact that is inevitable due to discretization errors.

Though conceptually straightforward, implementing contact forces in multibody fluid-structure simulations is far from simple in practice, with several competing possible strategies that are a topic of current research.  One approach is to use introduce a potential with short-range repulsion~\cite{}, but such methods often introduce stiffness. Alternatively, a space-time interference volume can guarantee a minimum separation between grains.  In a fluid context, this method was first introduced by Lu et al.~\cite{lu-rah-zor2017}, and extended to allow for much smaller separation distances of rigid bodies by the PI's previous graduate student~\cite{bys-sha-qua2019}. One Ph.D.~student, who will be jointly supervised by both PIs, will simulate sedimentation of eroding bodies by developing algorithms that resolve quadrature error and contact.

@@

\todo[inline]{Nick to Bryan: These changes you made look good to me. Did you want to add anything else, or is it okay to delete the above todo notes?}

@@

Transport through eroded media is a direct consequence of the pore
structure. Qualitative properties of an eroded media, such as channels
of high porosity can be modest in volume fraction, but still be
responsible for a large portion of the flux~\cite{Quaife2018}. 
\todo[inline]{There is no verb in the above sentence. Is there a word left out?}
To increase understanding of flow through eroded media, we will simulate
and statistically analyze transport through eroded geometries.

In an eroded geometry $\Omega$, the dimensionless
concentration $c$ is governed by
\begin{align}
  \pd{c}{t} + \uu \cdot \nabla c = \frac{1}{\Pe} \Delta c, 
  \label{eqn:advectionDiffusion}
\end{align}
where $\Pe$ is the dimensionless Peclet number. Here we assume that the
concentration $c$ is affected by the eroded geometry, but the geometry
is not altered further by the concentration. In the infinite Peclet
number limit, the concentration is governed by pure advection and the
characteristics of the concentration field are identical to the
streamlines (Figure~\ref{fig:100tracers}).
\begin{wrapfigure}[]{l}{0.4\textwidth}
  \includegraphics[width=0.4\textwidth]{figs/tort_eulerian100}
  \caption{\label{fig:100tortuosity} The tortuosity of an eroding
  geometry. The tortuosity can be computed using two different formulas
  (red vs.~blue marks). The red square is the tortuosity in
  Figure~\ref{fig:100tracers}. Their is qualitative agreement between
  our numerical simulations and a power-law model (black dashed
  line)~\cite{mat-kha-koz2008}.}
\end{wrapfigure}

The $j^{th}$ trajectories, or characteristics, is governed by
$\dot{\ss}_j(t) = \uu(\ss_j(t))$. Once these trajectories are computed,
using for example a high-order Runge-Kutta method, a statistical
analysis can be used to describe the transport. Two of the most important
measurements are the tortuosity and the anomalous dispersion, both which
are defined in terms of the length of the trajectories
\begin{align}
  \lambda_j(t) = \int_{0}^{t} \|\ss'(t')\| dt'.
\end{align}
Once $N_p$ trajectories are computed, the tortuosity is the
dimensionless number
\begin{align}
  T = \frac{1}{d} \left(\int_{S} u_1(x_0,y_0)\lambda(y_0) dy_0\, \right)
  \Bigg/ \left(\int_S u_1(x_0,y_0)\, dy_0 \right),
\end{align}
where $S$ is the vertical cross-section at $x=x_0$, $d$ is its length,
$u_1$ is the horizontal velocity, and $\lambda(y_0)$ is the length of
the trajectory starting at $(x_0,y_0)$ and ending at a vertical
cross-section at the other end of the channel. Relationships between the
tortuosity and porosity of a geometry have been developed by other
groups~\cite{kop-kat-tim1996, dud-koz-mat2011, mat-kha-koz2008}, and
preliminary results indicate that such models might be appropriate for
geometries whose porosity is increasing under the process of erosion
(Figure~\ref{fig:100tortuosity}).

@@

	More specifically, we plan to run large batches of the high-fidelity boundary-integral-equation-based numerical simulations of erosion and then leverage rapid advances in the field of machine learning and deep-neural networks to develop reduced models that parametrize the effects of erosion on the evolution of macroscopic medium properties. The output of the high-fidelity simulations can be transformed into essential an image format, enabling us to use cutting-edge advancements on deep neural-networks for image analysis \cite{pelt2018mixed}. Some important advantages of referenced work are the use of dilated convolutions in a deeply connected network, which first mixes scales and second reduces the number of super-parameters present and thus largely averts the risk of overfitting data. Both PIs will supervise a graduate student who will design and implement the deep-neural network for learning the essence of the high-fidelity simulations. Once the essential features are learned, they will be incorporated into reduced-order models that utilize the multiphase framework to enable more efficient and larger scale coarse-grained simulations of dynamic erosion.

As mentioned previously, researchers at FSU's Geophysical Fluid Dynamics Institute (GFDI) have investigated the mechanisms behind sinkhole formation with a series of scaled-down laboratory experiments~\cite{tao2014experimental}. Firstly, already existing data from these experiments offers a testbed for comparison against our newly developed simulations. Furthermore, we propose to run a set of new experiments at the facilities offered by GFDI, spearheaded by an undergraduate student in close collaboration with the graduate students and PIs. These new experiments will more deeply probe the mechanisms underlying sinkhole formation with the special design goal of comparison agains the numerical simulations. These experiments will be used to calibrate and validate our computations.
